Article Title

### 2014-01-27 plos.tex

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites} Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:03} and \ref{fig:05} is circled in blue. 
\\
    {\bf A.} The number of inferred latent units vs.~population size in each imaged site.
    {\bf B.} The average node degree for sparse partial correlations vs.~population size at each imaged site.
    {\bf C.} The average node degree in the sparse component in each site vs.~the number of latent units described by the low-rank component of the estimate.
    {\bf D.} The average sample correlations  vs.~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
    {\bf E.} Percentage overlap between the highest sample correlations and the  interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie 1-sparsity) of its sparse component. 
    {\bf F.} The percentage of negative interactions vs.~connectivity.
}
\label{fig:06}}
{\includegraphics{./figures/Figure06.pdf}}
\end{figure}

@@

Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity. For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.  Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .  
Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.   Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.

It is not yet clear which approach provides the best correspondence with anatomical connectivity. Little experimental evidence is available to answer this question.  The connectivity graphs inferred by various statistical methods are commonly reported as abstract representations of the functional connectivity without examining their relation to anatomical implications.  
Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.  However, the topological properties of functional connectivity graphs vary with the method of their inference \cite{Zalesky:2012}. Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.

Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the pairwise interactions on the activity of the other neurons, we must record from all neurons with which the pair interacts. Unobserved portions of the circuit can introduce false conditional dependencies between observed neurons. For this reason, the most statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}. In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the pairwise conditional dependencies. This may be the reason that partial correlations have not previously been used to describe  functional in cortical populations.

Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.  While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}. However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.  Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013}.  The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.  Recently, advanced imaging techniques have allowed for recordings from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.  These techniques may provide more incisive measurements of functional connectivity than classical electrophysiological recordings.

@@

\paragraph{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.  The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As  techniques improved to allow for recording from larger neuronal populations, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.  However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations. Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators. Regularization biases the solution toward a smaller subspace to counteract the effect of the sampling noise in the empirical data. However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.

@@

The partial correlation matrix with both sparse and low-rank components is computed from $C_{\sf sparse+latent}$ according to Eq.~\ref{eq:partial}; it includes the effects of interactions between the visible and latent units.  This estimate of the partial correlations was used in analyses of average partial correlations (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).  The sparse component $S$ was used separately in analyses of the connectivity graphs (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component alone without the low-rank component: 
\begin{equation}
P_{\sf sparse} = (S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
\end{equation}

>>>>

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites} Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:03} and \ref{fig:05} is circled in blue. 
\\
    {\bf A.} The number of inferred latent units vs.~population size in each imaged site.
    {\bf B.} The average node degree for sparse partial correlations vs.~population size at each imaged site.
    {\bf C.} The average node degree in the sparse component in each site vs.~the number of latent units described by the low-rank component of the estimate.
    {\bf D.} The average sample correlations  vs.~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
    {\bf E.} Percentage overlap between the highest sample correlations and the  interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie $1-sparsity$) of its sparse component. 
    {\bf F.} The percentage of negative interactions vs.~connectivity.
}
\label{fig:06}}
{\includegraphics{./figures/Figure06.pdf}}
\end{figure}

@@

Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity. For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.  Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.  Assuming that the population response follows a multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.   Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.

It is not yet clear which approach provides the best correspondence with anatomical connectivity. Little experimental evidence is available to answer this question.  The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to anatomy.  
Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.  However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}. Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.

Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the pairwise interactions on the activity of the other neurons, we must record from all neurons with which the pair interacts. Unobserved portions of the circuit can introduce false conditional dependencies between observed neurons. For this reason, statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}. In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of inference of the pairwise conditional dependencies. This may be the reason that partial correlations have not previously been used to describe  functional connectivity in cortical populations.

Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.  While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}. However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.  Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013}.  The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.  Recently, advanced imaging techniques have allowed for recordings from nearly every cell in a volume of cortical tissue  \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.  These techniques may provide more incisive measurements of functional connectivity than classical electrophysiological recordings.

@@

\paragraph{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies. The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As  techniques improved to allow for recording from larger neuronal populations, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}. However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations. Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators. Regularization biases the solution toward a small subspace to counteract the effect of  sampling noise in the empirical data. However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.

@@

The partial correlation matrix with both sparse and low-rank components is computed from $C_{\sf sparse+latent}$ according to Eq.~\ref{eq:partial}; it includes the effects of interactions between the visible and latent units.  This estimate of the partial correlations was used in analyses of average partial correlations (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).  The sparse component $S$ was used separately in analyses of the connectivity graphs (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component alone without the low-rank component: 
\begin{equation}
P_{\sf sparse} = -(S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
\end{equation}
